---
title: "Lab 02 - Plastic waste"
author: "Becky Upton"
date: !r format(Sys.Date())
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
# Load the tidyverse package

library(tidyverse)

```

```{r load-data, message=FALSE}

plastic_waste <- read.csv("data/plastic-waste.csv")

head(plastic_waste)

```

## Exercises

### Exercise 1

It looks like South America and Oceania have the least plastic waste per capita. It also looks like Africa and Asia may have the most plastic waste per capita but it isn't super clear based on these graphs.

```{r plastic-waste-continent}

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent) +
  geom_histogram()

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent) +
  geom_density(aes(y = after_stat(count)))

```

### Exercise 2

```{r plastic-waste-density}

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap, color = continent) +
  geom_density()

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap, color = continent, fill = continent) +
  geom_density(alpha = 0.2)

```
### Exercise 3

We put fill and color into aesthetic because we want those features to be mapped onto our variables. The alpha goes into the geom directly because we want the density portion itself to be transparent.

### Exercise 4

Remove this text, and add your answer for Exercise 4 here.

```{r plastic-waste-violin}
# insert code here
```

### Exercise 5

Remove this text, and add your answer for Exercise 5 here.

```{r plastic-waste-mismanaged}
# insert code here
```

### Exercise 6

Remove this text, and add your answer for Exercise 6 here.

```{r plastic-waste-mismanaged-continent}
# insert code here
```

### Exercise 7

Remove this text, and add your answer for Exercise 7 here.

```{r plastic-waste-population-total}
# insert code here
```

```{r plastic-waste-population-coastal}
# insert code here
```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

```
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
```

### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements
