---
title: "Class Notes"
output: word_document
---

Notes from lecture here:
https://bwiernik.github.io/progdata-class/r_scipts.html

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r blah}

sleep_times <- c(6, 9, 8, 4, 7, 6, 5, 8, 7, 8, 2.5)

mean(sleep_times)

```

c( ) stands for concatenate, it creates a vector or string of objects
  ### vectors must be all one type: numeric, logical (TRUE or FALSE), character, or factor
  ### a vector us a set of values that are all one type
  
?mean( )
###na.rm means remove N/A or missing values
  
  
```{r multiply}

minutes_sleep <- 60 * sleep_times

minutes_sleep

1:11 * sleep_times  ### multiplies each object in vector in sequential order

rep(1, 11) ### replicate function: first place is what it is replicating, second place is how many times it is repeating

```
  
```{r practice mean and other functions}

mean(sleep_times, 0.1)

mean(sleep_times, trim = 0.1) ### same thing but can remember what the heck we were doing

sd(sleep_times)

range(sleep_times)

```

```{r comparisons}

sleep_times < 6 ### how many people slept less than 6 hours?

sleep_times == 6

sleep_times != 6  ### ! means NOT, eg. ! TRUE means not true

sleep_times <= 6


which(sleep_times < 6) ### tells you which values in the sequence meet the criteria by giving the place spot of the value eg. value in place spot of 4 (4th value in sequence) meets the criteria

```
###doing multiple comparisons
& used for AND 
| used for OR

any()    ### set of logical vectors, is at least one of the values true?
all()    ### set of logical vectors, are all of the values true?


```{r compare specific stuff}

results <- c(TRUE, TRUE, FALSE)

correct <- c(TRUE, TRUE, FALSE)

results == correct

all(results == correct)

```

###subsetting
Use [] to subset

```{r subsetting}

sleep_times[1] ### gives first value in the vector

sleep_times[-1] ### gives all values except the first value

sleep_times[!is.na(sleep_times)] ### give me all values expect ones that are missing

na.omit(sleep_times) ### gives back vector without missing values

```

###modifying vectors

```{r}

sleep_times[2] <- NA  ### replaces second value in vector with NA

sleep_times[sleep_times < 5] <- NA ### replace all values less than five with NA

sleep_times

```

###data frames

```{r data frames}

nrow(mtcars)

mtcars$cyl ### pulls out cylinder column as a vector

mtcars[["cyl"]]  ### same ^

mtcars[[2]]  ### same ^

mean(mtcars[["mpg"]])

```

