---
title: "Lab 02 - Plastic waste"
author: "Becky Upton"
date: !r format(Sys.Date())
output: html_document
---

## Load packages and data

```{r setup, message=FALSE}
# Load the tidyverse package

library(tidyverse)

```

```{r load-data, message=FALSE}

plastic_waste <- read.csv("data/plastic-waste.csv")

head(plastic_waste)

```

## Exercises

### Exercise 1

It looks like South America and Oceania have the least plastic waste per capita. It also looks like Africa and Asia may have the most plastic waste per capita but it isn't super clear based on these graphs.

```{r plastic-waste-continent}

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent) +
  geom_histogram()

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap) +
  facet_wrap(~continent) +
  geom_density(aes(y = after_stat(count)))

```

### Exercise 2

```{r plastic-waste-density}

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap, color = continent) +
  geom_density()

ggplot(plastic_waste) +
  aes(x = plastic_waste_per_cap, color = continent, fill = continent) +
  geom_density(alpha = 0.2)

```
### Exercise 3

We put fill and color into aesthetic because we want those features to be mapped onto our variables. The alpha goes into the geom directly because we want the density portion itself to be transparent.

### Exercise 4

The density plot and individual data points show the variations between different countries within the continents that the box plot doesn't show.

```{r plastic-waste-violin}

ggplot(plastic_waste) +
  aes(x = continent, y =  plastic_waste_per_cap, fill = continent) +
  geom_boxplot()

ggplot(plastic_waste) +
  aes(x = continent, y =  plastic_waste_per_cap, fill = continent) +
  geom_boxplot() +
  geom_jitter() +
  ggdist::stat_slab(height = .3,
                    color = "black",
                    size = .2,
                    alpha = .5,
                    position = position_nudge(y = .5))
```

### Exercise 5

Just from looking at the graph, there doesn't seem to be much of a relationship between the variables. There may be a samll positive association between mismanaged plastic and plastic waste per capita.

```{r plastic-waste-mismanaged}

ggplot(plastic_waste) +
  aes(x = mismanaged_plastic_waste_per_cap, y =  plastic_waste_per_cap) +
  geom_point() 

```

### Exercise 6

It is very hard to tell if the association varies by continent from this graph. It looks like europe may have the least amount of outliers and less mismanaged plastic. It still seems that there may be a positive association between the variables for each continent.


```{r plastic-waste-mismanaged-continent}

ggplot(plastic_waste) +
  aes(x = mismanaged_plastic_waste_per_cap, y =  plastic_waste_per_cap, color = continent) +
  geom_point()

```

### Exercise 7

Neither of these variables seem to have a relationship with plastic waste per capita. Based on the ribbon of uncertainty around the trend line, the estimation of the association between coastal population and plastic waste per capita has less uncertainty than the estimation of the association between total population and plastic waste per capita. However, the uncertainty ribbons for both are huge.

```{r plastic-waste-population-total}

ggplot(plastic_waste) +
  aes(x = total_pop, y =  plastic_waste_per_cap) +
  geom_point() +
  geom_smooth()

```

```{r plastic-waste-population-coastal}

ggplot(plastic_waste) +
  aes(x = coastal_pop, y =  plastic_waste_per_cap) +
  geom_point() +
  geom_smooth()

```

### Exercise 8

Remove this text, and add your answer for Exercise 8 here.

```{r recreate-viz}
# insert code here
```

## Pro-Tips

### Excercise 3

Try this :D

```
ggplot(data = plastic_waste, 
       mapping = aes(x = continent, 
                     y = plastic_waste_per_cap)) +
  geom_violin()+
  geom_boxplot(width=.3, fill="green") +
  stat_summary(fun.y=median, geom="point") 
```

### Exercise 5 

Helpful reference:http://www.sthda.com/english/wiki/ggplot2-themes-and-background-colors-the-3-elements
