---
title: "Prog Data Lab 5"
author: "Becky Upton"
output: html_document
---

```{r set up}
library(tidyverse)
lotr  <- read_csv("https://raw.githubusercontent.com/jennybc/lotr-tidy/master/data/lotr_tidy.csv") |> 
  rename(Species = Race)
```

# Lab 5A
## 1. Would you say this data is in tidy format?

```{r look at data}
head(lotr)
```
Yes, this data is in tidy format.

## 2. Widen the data so that we see the words spoken by each species, by putting species as its own column.

```{r widen words}
lotr_wide <- lotr |> 
  pivot_wider(id_cols = c(-Species, -Words), 
              names_from  = Species, 
              values_from = Words)
```
## 3. Re-lengthen the wide LOTR data from Question 2 above.

```{r}
lotr_wide |> 
  pivot_longer(cols = c(Elf, Hobbit, Man), 
               names_to  = "Species", 
               values_to = "Words")
```
# Lab 5B
## 1. Reshape the dataset to a “longer” format. (Make each row 1 score. Have columns for family_id, family_member, wave, scale, and score.)

```{r, data 5B}
set.seed(123)
missing_w2_parent <- sample(1:500, 30)
missing_w2_child <- c(missing_w2_parent[1:5], sample(1:500, 25))
family  <- read_csv(
  "https://raw.githubusercontent.com/bwiernik/progdata/main/inst/tutorials/data/family_data.csv"
) |> 
  mutate(
    across(
      starts_with("w2") & contains("parent"),
      ~ ifelse(family_id %in% missing_w2_parent, NA_real_, .x)
    ),
    across(
      starts_with("w2") & contains("child"),
      ~ ifelse(family_id %in% missing_w2_child, NA_real_, .x)
    )
  )

family |> 
  DT::datatable(rownames = FALSE)
```

```{r, reshape}
family_longest <- family |>
  pivot_longer(
  cols = w1_comm_parent:w2_satis_child,
  names_to = c("wave", "measure", "family_member"), 
  names_sep = "_",
  values_to = "score"
) |> 
  mutate(wave = stringr::str_replace(wave, "w", "")) |> 
  mutate(wave = as.integer(wave))

print(family_longest)
```
## 2. Reshape the dataset to a “longer” format. Make each row 1 person. Have columns for family_id, family_member, wave, comm, and satis.
```{r, reshape again}
family_long <- family|> 
  pivot_longer(
    cols = -family_id,
    names_to = c("wave", ".value", "family_member"), 
    names_sep = "_"
  )

print(family_long)
```
## 3. Some families are missing wave 2 data for parent, child, or both. Which families are missing wave 2 data for at least one person?

```{r}
family_longest |>
  filter(wave == 2 & is.na(score))
```

